---
title: "Different Data Formats"
author: "Kumudini Bhave"
date: "October 23, 2016"
output:
  html_document:
    fontsize: 17pt
    highlight: pygments
    theme: cerulean
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---
     
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




********
     
# **Different Data Formats**
     
********
     
     
## Summary
     
 
We make an attempt to acquire data in different formats and use appropriate packages to acquire / load and process them into R readable format of data frame.
  
     
********

## R Code :
     
     
### Environment Setup

Loading Packages Used


```{r warning=FALSE, comment=FALSE, message=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=150)}
knitr::opts_chunk$set(message = FALSE, echo=TRUE)

# Library to read data file
library(RCurl)

library(knitr)

# For loading XML, JSON Data files
library(XML)
library(jsonlite)
library(htmltab)

# Library for data display in tabular format
library(DT)


```
     

********

### Different Data Formats

Data available in HTML, JSON and XML formats are loaded.

The data files considered are in following format. Below are the github links to the data files.


#### **Data Files**     

[HTML](https://raw.githubusercontent.com/DataDriven-MSDA/DATA607/master/Week7A/myfavbookshtml.html)
 
[JSON](https://raw.githubusercontent.com/DataDriven-MSDA/DATA607/master/Week7A/myfavbooksjson.json)


[XML](https://raw.githubusercontent.com/DataDriven-MSDA/DATA607/master/Week7A/myfavbooksxml.xml)


********

### HTML Data Acquisition

```{r  warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=80),echo=TRUE}


html.giturl <- "https://raw.githubusercontent.com/DataDriven-MSDA/DATA607/master/Week7A/myfavbookshtml.html"

# The htmltab() from package HTMLTAB directly gives data frame as output.

book.html <- htmltab::htmltab(html.giturl)

class(book.html)

# Verifying records and variables

nrow(book.html)
ncol(book.html)

# Renaming Columns
colnames(book.html) <- c("Title","Authors","Genre","Language","Publisher","Pages","Rating","Price(INR)")

# Display data frame content
datatable(book.html, rownames=FALSE)

```

********

### JSON Data Acquisition

```{r  warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=80),echo=TRUE}



json.giturl <- "https://raw.githubusercontent.com/DataDriven-MSDA/DATA607/master/Week7A/myfavbooksjson.json"

json.baseurlstr <- paste(readLines(json.giturl),collapse="")


book.jsondata <- fromJSON(json.baseurlstr)

class(book.jsondata)


# converting to Data Frame.

book.json <- as.data.frame(book.jsondata)

# Verifying records and variables
nrow(book.json)
ncol(book.json)

# Renaming Columns
colnames(book.json) <- c("Title","Authors","Genre","Language","Publisher","Pages","Rating","Price(INR)")

# Display data frame content
datatable(book.json, rownames=FALSE)


```


********

### XML Data Acquisition


```{r  warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=80),echo=TRUE}


xml.giturl <- "https://raw.githubusercontent.com/DataDriven-MSDA/DATA607/master/Week7A/myfavbooksxml.xml"

book.xmldata <- xmlParse(getURL(xml.giturl)) # get XML file contents

class(book.xmldata)
xmlSize(book.xmldata)

# Converting to data frame

book.xml <- xmlToDataFrame(book.xmldata)

# Verifying records and variables

nrow(book.xml)
ncol(book.xml)
head(book.xml)

# Renaming Columns

colnames(book.xml) <- c("Title","Authors","Genre","Language","Publisher","Pages","Rating","Price(INR)")

# Display data frame content
datatable(book.xml, rownames=FALSE)


```

********

The identical data frames are obtained although the source data formats were different.

Thus we observe that data can be acquired and loaded in different formats with usage of appropriate loading libraries and processing steps.


********     