---
title: "ReutersDocumentClassification"
author: "Kumudini Bhave"
date: "November 10, 2016"
output:
  html_document:
    fontsize: 17pt
    highlight: pygments
    theme: cerulean
    toc: yes
---

********


# **Reuters DataSet : Document Classification**

********


## Summary

**Performing Text Mining With Machine Learning Algorithm For Document Classification Of Reuters Dataset And Observing Performance Of Different Models For Classsification**


********



## R Code : Preparing Term Document Matrix From the Reuters DataSet


### Loading Packages Used


```{r warning=FALSE, comment=FALSE, message=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=150)}
knitr::opts_chunk$set(message = FALSE, echo=TRUE)



library("tm")
library("plyr")
library("class")
library("SnowballC")
library("stringr")
library("wordcloud")

library("RTextTools")
library("e1071")
library("tidyr")
library("DT")
library("ggplot2")



```


### Loading Dataset

Downloaded the Reuters Dataset from

Reuters-21578 DataSet can be downloaded as zip from : http://disi.unitn.it/moschitti/corpora.htm

Uploaded at
https://raw.githubusercontent.com/DataDriven-MSDA/DATA607/master/Week10A/Data

For this assignment, the zip was downloaded and extracted locally for different categories .
The Reuters dataset consists of 90 categories and classified into Test and Training as separate dataset for each category.

Here the categories of 'Trade' and 'MoneyFx' is used to perform document classification.


```{r  warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=80),echo=TRUE}



options(stringsAsFactors = FALSE)

datapath <- "F:/Data/Reuters21578_90Cat/training"
category <- c("trade","moneyfx")

# Trainig Data for Trade and MoneyFx categories

trade.directory <- "F:/Data/Reuters21578_90Cat/training/trade"
moneyfx.directory <- "F:/Data/Reuters21578_90Cat/training/moneyfx"    

# Test Data for Trade and MoneyFx categories

tradetest.dir <- "F:/Data/Reuters21578_90Cat/test/trade"
moneyfxtest.dir <- "F:/Data/Reuters21578_90Cat/test/moneyfx"

```

********

### Data Cleansing Function

The Corpus for 'Trade' and 'MoneyFx' categories need to be cleaned of punctuations, numbers and commong english language words. Doing so, facilitates the relevant terms to surface for text mining that would help build classification model.

```{r  warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=80),echo=TRUE}


cleanupDataCorpus <- function(dataCorpus){
     
     cleanDataCorpus <- tm_map(dataCorpus, removePunctuation)
     cleanDataCorpus <- tm_map(cleanDataCorpus, removeNumbers)
     
     cleanDataCorpus <- tm_map(cleanDataCorpus, str_replace_all, pattern
                              = "[[:punct:]]", replacement = " ")     
     cleanDataCorpus <- tm_map(cleanDataCorpus, tolower)
     cleanDataCorpus <- tm_map(cleanDataCorpus, removeWords, c('said','u.s',stopwords("english")))
     
     
     cleanDataCorpus <- tm_map(cleanDataCorpus, stemDocument)
     cleanDataCorpus<- tm_map(cleanDataCorpus, PlainTextDocument)
     return(cleanDataCorpus)
     
}

```


********

### Build Term Document Matrix For Training Data

```{r  warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=80),echo=TRUE}

trade.corpus <- Corpus(DirSource(directory=trade.directory, encoding = "ASCII"))
moneyfx.corpus <- Corpus(DirSource(directory=moneyfx.directory, encoding = "ASCII"))

trade.cleancorpus <- cleanupDataCorpus(trade.corpus)
moneyfx.cleancorpus <- cleanupDataCorpus(moneyfx.corpus)

trade.tdm <- TermDocumentMatrix(trade.cleancorpus)
moneyfx.tdm <- TermDocumentMatrix(moneyfx.cleancorpus)

trade.tdm <- removeSparseTerms(trade.tdm, 0.7)
moneyfx.tdm<- removeSparseTerms(moneyfx.tdm, 0.7)

trade.datamatx <- t(data.matrix(trade.tdm))
moneyfx.datamatx <- t(data.matrix(moneyfx.tdm))

head(trade.datamatx)
head(moneyfx.datamatx)


trade.df <- as.data.frame(trade.datamatx)
moneyfx.df <- as.data.frame(moneyfx.datamatx)


```

********


### Build Term Document Matrix For Test Data

```{r  warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=80),echo=TRUE}

tradetest.corpus <- Corpus(DirSource(directory=tradetest.dir, encoding = "ASCII"))
moneyfxtest.corpus <- Corpus(DirSource(directory=moneyfxtest.dir, encoding = "ASCII"))

tradetest.cleancorpus <- cleanupDataCorpus(tradetest.corpus)
moneyfxtest.cleancorpus <- cleanupDataCorpus(moneyfxtest.corpus)

tradetest.tdm <- TermDocumentMatrix(tradetest.cleancorpus)
moneyfxtest.tdm <- TermDocumentMatrix(moneyfxtest.cleancorpus)

tradetest.tdm <- removeSparseTerms(tradetest.tdm, 0.7)
moneyfxtest.tdm<- removeSparseTerms(moneyfxtest.tdm, 0.7)

tradetest.datamatx <- t(data.matrix(tradetest.tdm))
moneyfxtest.datamatx <- t(data.matrix(moneyfxtest.tdm))

head(tradetest.datamatx)
head(moneyfxtest.datamatx)


tradetest.df <- as.data.frame(tradetest.datamatx)
moneyfxtest.df <- as.data.frame(moneyfxtest.datamatx)

```



********


### Aligning Test Data Terms with Training Data Terms


Filtering for columns present in test and training and then adding the columns/terms present only in training.
Attempt to keep the same terms in test and training 


#### subset trade testing data by colnames (ie. terms) from training trade data

```{r  warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=80),echo=TRUE}

trade.df.mat <- as.matrix(trade.df)
tradetest.df.mat <-  as.matrix(tradetest.df)


tradetestfiltered1 <- data.frame(tradetest.df.mat[,intersect(colnames(tradetest.df.mat),
                                           colnames(trade.df.mat))])

tradetestfiltered2 <- read.table(textConnection(""), col.names = colnames(trade.df.mat))


tradetestfiltered <- rbind.fill(tradetestfiltered1, tradetestfiltered2)

ncol(tradetestfiltered)

```



#### subset moneyfx testing data by colnames (ie. terms) from moneyfx training data

```{r  warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=80),echo=TRUE}


moneyfx.df.mat <- as.matrix(moneyfx.df)
moneyfxtest.df.mat <-  as.matrix(moneyfxtest.df)


moneyfxtestfiltered1 <- data.frame(moneyfxtest.df.mat[,intersect(colnames(moneyfxtest.df.mat),
                                                             colnames(moneyfx.df.mat))])

moneyfxtestfiltered2 <- read.table(textConnection(""), col.names = colnames(moneyfx.df.mat))


moneyfxtestfiltered <- rbind.fill(moneyfxtestfiltered1, moneyfxtestfiltered2)

ncol(moneyfxtestfiltered)

```


********



### Binding 'Category' Column To The Data Frame

```{r  warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=80),echo=TRUE}


trade.df <- cbind(trade.df, category= rep("trade"))
moneyfx.df <- cbind(moneyfx.df, category = rep("moneyfx"))

tradetestfiltered <- cbind(tradetestfiltered, category= rep("trade"))
moneyfxtestfiltered <- cbind(moneyfxtestfiltered, category = rep("moneyfx"))

```


********

### Stacking different categories TDM together

We now have two dataframes  , one each category , with the content as rows and the terms as variables / columns, and a category column appended to denote the category.

We now stack both the tdms 


#### Stacking the TDMS of different categories for training data

```{r  warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=80),echo=TRUE}


tdm.stackedcategory <- rbind.fill(trade.df, moneyfx.df)

nrow(trade.df)
nrow(moneyfx.df)
nrow(tdm.stackedcategory)

tdm.stackedcategory [is.na(tdm.stackedcategory)] <- 0

ncol(trade.df)
ncol(moneyfx.df)
ncol(tdm.stackedcategory)

```


#### Stacking the TDMS of different categories for test data


```{r warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=80),echo=TRUE}

tdm.teststackedcategory <- rbind.fill(tradetestfiltered, moneyfxtestfiltered)

nrow(tradetestfiltered)
nrow(moneyfxtestfiltered)
nrow(tdm.teststackedcategory)

tdm.teststackedcategory [is.na(tdm.teststackedcategory)] <- 0

ncol(tradetestfiltered)
ncol(moneyfxtestfiltered)
ncol(tdm.teststackedcategory) 

```


We now see the number of total rows in stacked training TDM and stacked test TDM is equivalent to the sum of total rows for each category TDM.

********
********

### Word Cloud For Training Sets For Trade And Moneyfx Categories

findAssocs(tdm.stackedcategory, terms = "billion", corlimit = 0.3)

**Trade Word Cloud**

We find that words like 'trade','billion', 'year' have been more prominently used
```{r  warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=80),echo=TRUE}

# Word Cloud for Trade

wordcloud(trade.cleancorpus, max.words =20 , random.order = FALSE, colors = brewer.pal(8,"Dark2"),  scale = c(5,0.5), vfont = c("gothic english","plain"))


```


**Moneyyfx Word Cloud**

We find that words like 'bank','market', 'pct' 'Ddollar' have been more prominently used.
```{r  warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=80),echo=TRUE}

# Word Cloud for Moneyfx

wordcloud(moneyfx.cleancorpus,max.words =20 , random.order = FALSE,colors = brewer.pal(8,"Dark2") ,  scale = c(5,0.5), vfont=c("serif","plain"))



```


********

## R Code : Preparing Data For Model Based Classification


Here we separate the training data and test data both from the category column so as to provide to train the model and compare it later for accuracy.


```{r  warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=80),echo=TRUE}

# training data with only category column specified
tdm.traincategorycol <- tdm.stackedcategory[,"category"]

# training data without category column specified
tdm.train <- tdm.stackedcategory[,!colnames(tdm.stackedcategory) %in% "category"]

# test data with only category column specified
tdm.testcategorycol <- tdm.teststackedcategory[,"category"]

# test data without category column specified
tdm.test <- tdm.teststackedcategory[,!colnames(tdm.teststackedcategory) %in% "category"]


```

********

###  MODEL KNN (K Nearest Neighbour)

```{r  warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=80),echo=TRUE}

knn.pred <- knn(tdm.train, tdm.test, tdm.traincategorycol)



# Accuracy Test

confusion_matx <- table("Predictions" = knn.pred, Actual = tdm.testcategorycol)
confusion_matx

# Calculating Accuracy
accuracy_1 <- sum(diag(confusion_matx))/ nrow(tdm.test) 
accuracy_1

# Another way to calculate accuracy

true_moneyfx <- 179
true_trade <- 112

false_moneyfx <- 5
false_trade <- 0
     
     
total <- true_moneyfx + true_trade + false_moneyfx + false_trade
accuracy_2 <- (true_moneyfx + true_trade) / total 
accuracy_2
error <- 1 - accuracy_2
error

```


********


```{r  warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=80),echo=TRUE}


container <- create_container(tdm.train, tdm.stackedcategory$category, trainSize=1:907, virgin=FALSE)

predtestcontainer <- create_container(tdm.test, labels= 0:296, testSize = 1:296, virgin=FALSE)

```



### MODEL SVM (Support Vector Machine)

```{r  warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=80),echo=TRUE}

SVM <- train_model(container,"SVM")
results.SVM <- classify_model(predtestcontainer, SVM)
datatable(results.SVM)


```

********

### MODEL MAXENT (Max Entropy)

```{r  warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=80),echo=TRUE}

MAXENT <- train_model(container,"MAXENT")
results.MAXENT <- classify_model(predtestcontainer, MAXENT)
datatable(results.MAXENT)

```

********

### MODEL BAGGING (Bag)

```{r  warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=80),echo=TRUE}

BAGGING <- train_model(container,"BAGGING")
results.BAGGING <- classify_model(predtestcontainer, BAGGING)
datatable(results.BAGGING)

```

********

### MODEL Random Forest

```{r  warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=80),echo=TRUE}


RF <- train_model(container,"RF")
results.RF <- classify_model(predtestcontainer, RF)
datatable(results.RF)

```


********

### MODEL TREE

```{r  warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=80),echo=TRUE}

TREE <- train_model(container,"TREE")
results.TREE <- classify_model(predtestcontainer, TREE)
datatable(results.TREE)

```

********

### Performance Comparison Of Different Models

```{r  warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=80),echo=TRUE}


cat_check <- data.frame(
     correct_cat = tdm.testcategorycol[1:296],
     svm = as.character(results.SVM[,1]),
     maxent = as.character(results.MAXENT[,1]),
     bagging = as.character(results.BAGGING[,1]),
     rf = as.character(results.RF[,1]),
     tree = as.character(results.TREE[,1]),
     stringAsFactors = F)


# SVM Performance

svm_table <- table(cat_check[,1] == cat_check[,2])
addmargins(svm_table)
psvm <- prop.table(svm_table)
psvm <-as.data.frame(psvm)
psvm$Model <- "SVM"


# MAXENT Performance

maxent_table <- table(cat_check[,1] == cat_check[,3])
addmargins(maxent_table)
pmaxent <- prop.table(maxent_table)
pmaxent <- as.data.frame(pmaxent)
pmaxent$Model <- "MAXENT"


# Bagging Performance

bagging_table <- table(cat_check[,1] == cat_check[,4])
addmargins(bagging_table)
pbag <- prop.table(bagging_table)
pbag <-as.data.frame(pbag)
pbag$Model <- "BAGGING"


# RF Performanace

rf_table <- table(cat_check[,1] == cat_check[,5])
addmargins(rf_table)
prf <- prop.table(rf_table)
prf <- as.data.frame(prf)
prf$Model <- "RF"


# Tree Performanace

tree_table <- table(cat_check[,1] == cat_check[,6])
addmargins(tree_table)
ptree <- prop.table(tree_table)
ptree <-as.data.frame(ptree)
ptree$Model <- "TREE"


```




```{r  warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=80),echo=TRUE}



pmodels <- rbind.fill(as.data.frame(psvm)
,as.data.frame(pmaxent)
,as.data.frame(pbag)
,as.data.frame(prf),as.data.frame(ptree)
)




colnames(pmodels) <- c("Status" , "Frequency", "Model")

pmodels <- rbind(pmodels, c('FALSE',error,"KNN"),c('TRUE',accuracy_2,"KNN"))


datatable(pmodels)

```

********

### Plotting Performance Of Models

```{r  warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=80),echo=TRUE}


ggplot(pmodels, aes(x=Model, y=Frequency)) + ggtitle("Model Performance Plot") + geom_point(aes(colour = Status), size = 4) + theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

********

### Conclusion

We find that Random Forest Model and K-Nearest Neighborr gives optimum performance in classifying the documents correctly with 99.32% and 98.31% accuracy respectively.
Followed by Support Vector Machine 97.97% and Bagging 97.29% accuracy
Hence Randome Forest would be the best mdel for classifying further test documents dataset for 'Trade' and 'Moneyfx' categories , based on saved models.



********









