---
title: "Migration From Relational DB to NoSQL DB"
author: "Kumudini Bhave"
date: "November 15, 2016"
output: 
     html_document:
          depth : 5
          theme: united
          highlight: tango
---


*********


## **Migration to NoSQL Databases** {.tabset}
 
 Loading Packages Used

```{r warning=FALSE, comment=FALSE, message=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=150)}
knitr::opts_chunk$set(message = FALSE, echo=TRUE)

# Libraries Used : DT , RMySQL, dplyr

# Library for data display in tabular format
library(DT)
# Library for db operations
library(RMySQL)
# Library for data structure operations
library(dplyr)
library(jsonlite)
library(mongolite)

```



*********


### MongoDB Installation


MongoDb (64 bit version) was downloaded and installed in default directory under "C:\\Program Files"
The new mongolite package was used for client connectivity from R to MongoDB database.


To set up the MongoDB environment:

MongoDB requires a data directory to store all data. MongoDB's default data directory path is \\data\\db.
Hence directory for data and log is created

"F:\\mongodb\\data"
"F:\\mongodb\\log"

The mongodb startup requires 


The following command 
"C:\\Program Files\\MongoDB\\Server\\3.2\\bin\\mongod.exe" --dbpath F:\\mongodb\\data


To connect to MongoDB:

One needs to start the mongo.exe shell in another Command Prompt.

"C:\\Program Files\\MongoDB\\Server\\3.2\\bin\\mongo.exe"

The default database is 'test'

In mongo sheel one can connect to test using following command at mongo shell prompt :

>use test

One can check the collections ( equivalent to tables concept from relational databases) using :

>show collections

The default collection available in 'test' database is 'restaurants'

*********

#### References:

The MongoDB Manual : (https://docs.mongodb.com/manual/crud/

Tutorial Mongolite : https://cran.r-project.org/web/packages/mongolite/mongolite.pdf

*********
 
### MongoDB : Connection, Populating

*********

A connection to MYSQL relational databse is done and the data from 'entertainment' schema and 'tb' schema are extracted to perform a migration to NoSQL database of MongoDB.

*********

#### MYSQL Connection

```{r  tidy=TRUE, tidy.opts=list(width.cutoff=80),echo=TRUE}

mysql_dbconn <- function(schema, uid, pwd){
  
     rmysql.settingsfile<-"C:/ProgramData/MySQL/MySQL Server 5.6/my.ini"
     conn <- dbConnect(RMySQL::MySQL(), dbname = schema, username=uid ,password = pwd)
     return(conn)          
}

```


*********

#### Forming MYSQL DB Connection


```{r  tidy=TRUE, tidy.opts=list(width.cutoff=80),echo=TRUE}


conntb <- mysql_dbconn("tb","root","WhiteLotus21")

# List the database tables in entertainment schema
entdbtables <- dbListTables(conntb)

```


*********

#### Capturing tb database and entertainment database from MYSQL 

Loading tb and movie data in R

```{r tidy=TRUE, tidy.opts=list(width.cutoff=80), echo=TRUE}


if(isIdCurrent(conntb))
{
     tbdb <-  dbGetQuery(conntb,"select country, year, sex, child, adult, elderly from tb where tb.year > 2010")
     
}
    

connent <- mysql_dbconn("entertainment","root","WhiteLotus21")

if(isIdCurrent(conntb))
{ 
     moviesdb <-  dbGetQuery(connent,"select movieid, moviename, genre, yearreleased, leadcast, dirrectedby, countryreleased, language, watchtimemin, certificate, awards, info from movies, genre, certificate where movies.genre = genre.genreid and movies.certificate = certificate.certificateid")
     
      genredb <-  dbGetQuery(connent,"select genreid, genretype from genre")

      certificatedb <- dbGetQuery(connent,"select certificateid, certificatedesc from certificate")
     
      surveydb <- dbGetQuery(connent,"select respondentid, movieid, rating, responsedt from survey")
      
     
  
}
```


*********



#### MongoDB : Forming Connection, Checking For Existin "Restaurants" Collection in 'test' db

```{r  tidy=TRUE, tidy.opts=list(width.cutoff=80),echo=TRUE}

c = mongo(collection = "restaurants", db = "test")
c$count()

```


#### MongoDB : Checking For Existing "tb" Collection in 'tbdatabase' db


```{r  tidy=TRUE, tidy.opts=list(width.cutoff=80),echo=TRUE}
# To take care of the insertion of 'tb' data during the re-run of this Rmd , we drop it initially if existing at all



tb = mongo(collection = "tb", db = "tbdatabase")
tb$count()

tb$insert(tbdb)
tb$count()


```


```{r  tidy=TRUE, tidy.opts=list(width.cutoff=80),echo=TRUE}
# To take care of the insertion of 'tb' data during the re-run of this Rmd , we drop it at the end.

tb$drop()

```

#### MongoDB : Checking For Existing "entertainment" Collection in 'entertainment' db


```{r  tidy=TRUE, tidy.opts=list(width.cutoff=80),echo=TRUE}

mov = mongo(collection = "movies", db = "entertainment")
gen = mongo(collection = "genre", db = "entertainment")
cert = mongo(collection = "certificate", db = "entertainment")

survey = mongo(collection = "survey", db = "entertainment")

```



```{r  tidy=TRUE, tidy.opts=list(width.cutoff=80),echo=TRUE}
# To take care of the insertion of all collections in 'entertainment' database during the re-run of this Rmd , we drop it at the end.


survey$drop()
gen$drop()
cert$drop()
mov$drop()
```


#### Populating the dataframes from MYSQL to MongoDB

```{r  tidy=TRUE, tidy.opts=list(width.cutoff=80),echo=TRUE}

mov$count()
mov$insert(moviesdb)


gen$count()
gen$insert(genredb)


cert$count()
cert$insert(certificatedb)



survey$count()
survey$insert(surveydb)





mov$count()
gen$count()
cert$count()
survey$count()

```

*********

### MongoDB : Data Retrieval

********* 

#### Basic Data Retrieval

*********

```{r  tidy=TRUE, tidy.opts=list(width.cutoff=80),echo=TRUE}



# Selecting distinct respondents from Survey collection

survey$distinct("respondentid")


# Select respondent ids who did ot respond/ rate to MN1 movie code, ie Ice Age

survey$distinct("respondentid", '{"movieid" : "MN1", "rating": null }')

# Select respondent ids who respond/ rate to MN1 movie code, ie Ice Age , handle null

survey$find('{"movieid" : "MN1"}', handler = NULL)




# Add index and sort By Rating

survey$index(add="rating")

mysurveydata <- survey$find(sort='{"rating":-1}')
datatable(mysurveydata)

```


********

#### Find Movie And Average Rating


```{r  tidy=TRUE, tidy.opts=list(width.cutoff=80),echo=TRUE}

survey$aggregate( '[{"$group":{"_id" : "$movieid","avgRating" : {"$avg" : "$rating"}}}]')


# Logical And

western2016 <- mov$find('{"genre":"W","yearreleased":2016}')

datatable(western2016)

# Movie Either Animated OR Action , Sorted By Movie Name Descending

actionanimation <- mov$find('{"genre": { "$in" : ["N","A"]}}',fields = '{"moviename":1,"leadcast":1,"info":1} ', sort='{"moviename" : -1}')

datatable(actionanimation)

 
# Movies (Either Western OR Horror) And Released In 2016
 
westernorhorror2016 <-  mov$find('{"genre": { "$in" : ["W","H"]} , "yearreleased" : 2016 }',fields = '{"moviename":1,"leadcast":1,"info":1}')



datatable(westernorhorror2016)

# The collection can be stored to file and restored back

# Dump to bson
dump <- tempfile()
mov$export(file(dump), bson = TRUE)

# Remove the collection
mov$drop()

# Restore
mov$count()
mov$import(file(dump), bson = TRUE)
file(dump)
mov$count()


# Attempt to insert Json file in MongoDB

#donut <- mongo("zips", verbose=FALSE)


#donutdf <- stream_in(url("https://raw.githubusercontent.com/DataDriven-MSDA/DATA607/master/Week12P4/donut.json"))
                           
#donut$insert(donutdf)

#Alternatively

#stream_in(url("https://raw.githubusercontent.com/DataDriven-MSDA/DATA607/master/Week12P4/donut.json"),handler=function(df){donut$insert(df)})


# Could not import, This gives an error premature EOF

```


*********

### MongoDB (NoSQL) Compared To Relational (MYSQL / Oracle)

*********

MongoDB is an open source document-oriented NoSQL database which stores data in the form of JSON-like objects.
Hence a collection in MongoDB is equivalent to Table in relational database
And a document in a collection in MongoDB is equivalent to a row in relational database table.

NoSQL and relational databses differ in the manner they represent data.
While a row stores data in columns, document stores in JSON like structure.

MongoDB documents support dynamic schema, i.e. documents in same collection can have different schema , one can have 5 fileds , while other can have more or less than five.
This allows complex hierarchical /tree structured schemas is be efficiently stored in MongoDB.

Relationships in relational databse can be achieved through primary key /foreign key constraints ; No such mapping is present in MongoDB other than linking / emnbedding of documents .

However the document type representation and the unconventional way of querying seems to be bit hard to get used to syntactically.
The plain english like SQL of relational database seems more easy to adapt to.

*********
