---
title: "Data Transformation Project"
author: "Kumudini Bhave"
date: "October 23, 2016"
output:
  html_document:
    fontsize: 17pt
    highlight: pygments
    theme: cerulean
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---
     
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




********
     
# **Data Transformation :  Using *tidyr* and *dplyr* Packages**
     
********
     
     
## Summary
     
 
We make an attempt to acquire different datasets, process them into R readable format of data frame and perform data transformation in long /wide formats using tidyr / dplyr / reshape  packages.
  
     
********

## R Code :
     
     
### Environment Setup

Loading Packages Used


```{r warning=FALSE, comment=FALSE, message=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=150)}
knitr::opts_chunk$set(message = FALSE, echo=TRUE)

# Library to read data file
library(RCurl)
library(stringr)
library(knitr)
     
library(ggplot2)

# Library for data display in tabular format
library(DT)
library(tidyr)
library(dplyr)
library(reshape)
```
     

********

### Different Data Formats

Data available in HTML, JSON and XML formats are loaded.

The data files considered are in following format. Below are the github links to the data files.


********

### Data Set 1 : Birth DataSet {.tabset}


#### Loading Data /Cleansing


```{r  warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=80),echo=TRUE}



birthdata.giturl <- "https://raw.githubusercontent.com/DataDriven-MSDA/DATA607/master/Week6P2/birthdata.csv"


birthdata.gitdata <- getURL(birthdata.giturl)


birthdata <- read.csv2(text=birthdata.gitdata,                      header=T,sep=",",dec=".",stringsAsFactors= FALSE)

# Verifying records and variables

nrow(birthdata)
ncol(birthdata)
head(birthdata)



# Renaming Columns
colnames(birthdata) <- c("Case","MotherAge","BabyYear_1","BabyYear_2","BabyYear_3","BabyWt_1","BabyWt_2","BabyWt_3")

# Display data frame content
datatable(birthdata, rownames=FALSE)

```


#### Data Transformation

Using reshape package

The birth data is available on wide format.
Converting to long with relevant columns , Using reshape() to convert to long format

Reshape() provides the best way to solve this kind of dataset.
The column names need to follow a proper format to convert to long.

```{r  tidy=TRUE, tidy.opts=list(width.cutoff=80),echo=TRUE}



birthdatalong <- reshape(birthdata, varying=c("BabyYear_1","BabyYear_2","BabyYear_3", "BabyWt_1", "BabyWt_2", "BabyWt_3"), direction ="long", idvar="Case", sep="_")


datatable(birthdatalong, rownames=FALSE)


```

#### Analyzing Birth DataSet

```{r  tidy=TRUE, tidy.opts=list(width.cutoff=80),echo=TRUE}

# Average baby Weight ForEeach Mother

     birthdata_meanbabywt <- birthdatalong %>% select( Case, BabyWt) %>% group_by(Case) %>%  dplyr::summarise(AvgBabyWt = mean(BabyWt, na.rm=TRUE))

birthdata_meanbabywt


# Average Mother Age For Each BabyBirth

     birthdata_meanbabywtage <- birthdatalong %>% select( time, MotherAge) %>% group_by(time) %>%  dplyr::summarise(AvgMotherAgeFirstBaby = mean(birthdatalong$MotherAge, na.rm=TRUE)) 


library(scales)     
     
ggplot(birthdatalong, aes(y=birthdatalong$BabyWt, x=birthdatalong$MotherAge)) + geom_point(aes(color=factor(Case), group=Case)) + scale_color_discrete(name="Case") + labs(title = "Mother Age Vs Baby Weight ", x= "Mother Age", y="Baby Weight") 



```    


Since there are many NA values , and also the point may be over lapping in the plot.




********

### Data Set 2 : CitiBike {.tabset}

Courtesy : Dataset from Sharon Morris

#### Loading Data


```{r  warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=80),echo=TRUE}



bike.giturl <- "https://raw.githubusercontent.com/DataDriven-MSDA/DATA607/master/Week6P2/201606-citibike-tripdata.csv"


bike.gitdata <- getURL(bike.giturl)


bikedata <- read.csv2(text=bike.gitdata,                      header=T,sep=",",dec=".",stringsAsFactors= FALSE)

# Verifying records and variables

nrow(bikedata)
ncol(bikedata)
datatable(head(bikedata))

###############################################################################

```


#### Data Cleansing

Lets select only the columns that  we need to focus on for gender based trip duration between stations

```{r  warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=80),echo=TRUE}

head(bikedata)

colnames(bikedata) <- c("tripduration","starttime","stoptime","startid","startname","startlat","startlong","stopid","stopname","stoplat","stoplong","bikeid","usertype","birthyear","gender")


bikedata <- bikedata %>% dplyr::select(tripduration,starttime,stoptime,startid,stopid,usertype,gender)


###############################################################################   
```
   
   
    
#### Data Transformation/ Analyzing

          
```{r  warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=80),echo=TRUE}

# Since the data is too large we filter the same. Choosing trip duration(since it is in seconds) for greater than 5 hours
          
bikedatasub_18k <- bikedata %>% dplyr::filter( tripduration >= 18000) 

# Adding Column of tripduration in hours

bikedatasub_triphr <- bikedatasub_18k %>% mutate(round(triphours = tripduration /3600))

     
# caclulate mean time taken between Stations :Lexington Ave And Pearl St from base bike data

bikedatasub2 <- bikedata %>%  filter(startid == "120" & stopid == "82") %>% mutate(tripname = "LexPearl") %>%   summarize(Meantime=mean(tripduration, na.rm = TRUE))

bikedatasub2

# Finding total number of different trips in  bikedata set for above 5 hours and the mean time for each trip and arrange descending mean trip duration

bikedatasub_tripsnmean <- bikedatasub_18k %>%
  group_by(startid,stopid) %>% summarize(total = n(), tripmeantime = mean(tripduration, na.rm = TRUE))  %>%  arrange(desc(tripmeantime))


datatable(bikedatasub_tripsnmean)

# grouping by usertype

bikedatasub3<- bikedatasub_18k %>%  group_by(usertype) %>%  dplyr::summarise(avgdurn = mean(bikedatasub_18k$tripduration, totaltrip = n()))


###############################################################################
```


********

### Data Set 3 : Religion And Income  {.tabset}


(Courtesy : Dataset from Marco Sequeira)

#### Loading Data

```{r  warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=80),echo=TRUE}


relinc.giturl <- "https://raw.githubusercontent.com/DataDriven-MSDA/DATA607/master/Week6P2/religionincome.csv"


relinc.gitdata <- getURL(relinc.giturl)


relincdata <- read.csv2(text=relinc.gitdata, header=T,sep=",",dec=".",stringsAsFactors= FALSE)

# Verifying records and variables

nrow(relincdata)
ncol(relincdata)
head(relincdata)

###############################################################################
```


#### Data Cleansing

```{r  warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=80),echo=TRUE}


# Renaming Columns

colnames(relincdata) <- c("Tradition","ForLessThan30","From30To49","From50To99","MoreThan100","Sample")


# Removing % sign from the Columns

relincdata$ForLessThan30 <- as.numeric(str_extract(relincdata$ForLessThan30,"[[:digit:]]{2,}"))

relincdata$From30To49 <- as.numeric(str_extract(relincdata$From30To49,"[[:digit:]]{2,}"))

relincdata$From50To99 <- as.numeric(str_extract(relincdata$From50To99,"[[:digit:]]{2,}"))

relincdata$MoreThan100 <- as.numeric(str_extract(relincdata$MoreThan100,"[[:digit:]]{2,}"))

# Removing , from Sample

relincdata$Sample <- as.numeric(gsub(",","", relincdata$Sample))

# Display data frame content
datatable(relincdata, rownames=FALSE)



###############################################################################
```

********

####  Data Transformation/ Analysis

Using tidyr package


Using gather() to convert to long format


```{r  tidy=TRUE, tidy.opts=list(width.cutoff=80),echo=TRUE}

tidyrelincdata <- relincdata %>% 
     gather( "Category", "Pct", 2 :5, na.rm=FALSE)


tidyrelincdata$PctVal =round(((as.numeric(tidyrelincdata$Pct )/ 100)  * as.numeric(tidyrelincdata$Sample)) , 2)

datatable(tidyrelincdata)


  G <- ggplot(tidyrelincdata, aes(x=tidyrelincdata$Category,y=tidyrelincdata$PctVal)) + geom_point(aes(color=Tradition)) + ggtitle("Religion And Income Categoris") + xlab("Category") +  ylab("People In Category")
    
  G

```     


From the plot we notice that Evangelist Protestant followers are the maximum working in less than 30K category as well as in the 50k - 99k category

While themore than 100 K category is led by Unaffiliated followers.

The least in each category are Orthodox Christian followers.


********     